### 2017-01-07 tikz/graphbothCostsUpd.tex

\begin{tikzpicture}
  \node[node] (A) {};
  \node[below=0.5 of A] (d1) {};
  \node[node, below=0.5 of d1] (B) {};
  \node[right=1 of A] (d2) {};
  \node[node, right=1 of d2] (D) {};
  \node[node] (C) at (d1-|d2) {};
  \node[node] (E) at (B-|D) {};
  \node[node, above=0.5 of d2] (F) {};

  \draw[edge, p1] (A) to node[right, black, label={[black,label distance=-0.2cm]right:\tiny$8$}] {\tiny\bcancel{$7$}} (B);
  \draw[edge, both, dotted] (A) to node[above, black] {\tiny$8$} (C);
  \draw[edge, p1] (B) to node[below, black] {\tiny$2$} (C);
  \draw[edge, both] (C) to node[above, black] {\tiny$3$} (D);
  \draw[edge, p1] (C) to node[below, black] {\tiny$6$} (E);
  \draw[edge, p2] (A) to node[above, black] {\tiny$5$} (F);
  \draw[edge, p2] (D) to node[above, black] {\tiny$2$} (F);
\end{tikzpicture}