### 2017-01-07 tikz/graphp1ContDel.tex

\begin{tikzpicture}
  \node[node] (A) {};
  \node[below=0.5 of A] (d1) {};
  \node[node, below=0.5 of d1] (B) {};
  \node[right=1 of A] (d2) {};
  \node[node, right=1 of d2] (D) {};
  \node[node] (C) at (d1-|d2) {};
  \node[node] (E) at (B-|D) {};

  \node[fit=(A) (B), draw, rectangle, rounded corners, black] (new) {};

  \draw[edge, p1, bend right, dotted] (C) to (new.east);
  \draw[edge, p1, bend left] (C) to (new.east);
  \draw[edge, p1] (A) to (B);
  \draw[edge, p1] (C) to (D);
  \draw[edge, p1] (C) to (E);
\end{tikzpicture}