### 2017-01-11 slides.tex
added lemma and theorem for non-matroid space

\documentclass{beamer}
\usepackage{pgfplots}
\usepackage{tikz}
\usepackage{csquotes}
\usepackage{cancel}
\usetheme{metropolis}
\title{Resource Buying Games}
\author[C. Welzel]{Christoph Welzel}

\newcommand{\tupel}[1]{\left(#1\right)}
\newcommand{\set}[1]{\left\{#1\right\}}

\DeclareMathOperator{\ex}{ex}

\begin{document}
\input{tikz/settings}
\maketitle
\section{Introduction}
\begin{frame}
\end{frame}

\subsection{Game definition}
\begin{frame}
  \frametitle{Definition: congestion model}
  \begin{equation*}
    \mathcal{M} = \tupel{N, E, \mathcal{S}, \left(d_{i}\right)_{i\in N},
    \left(c_{e}\right)_{e\in E}}
  \end{equation*}
  \vspace{-1cm}
  \begin{itemize}
    \item $N$: set of players
    \item $E$: set of resources
    \item $\mathcal{S} = \times_{i\in N}\mathcal{S}_{i}$ with $\mathcal{S}_{i}$
      set of desired resources $S_{i}\subseteq E$: set of configurations
    \item $d_{i}$: demand of player $i$ for her resource
      \begin{itemize}
        \item $d_{i} = 1$ for an unweighted game
        \item $\ell_{e}(S) = \sum_{i\in N:e\in S_{i}}d_{i}$: load of $e$
          in configuration $S\in\mathcal{S}$
      \end{itemize}
    \item $c_{e}:\mathbb{N}\rightarrow\mathbb{N}$ with
      $c_{e}(S) = c_{e}(\ell_{e}(S))$: cost for the resource $e$ under the load
      of $S$
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Definition: weighted resource buying game}
  \begin{equation*}
    \mathcal{M} = \tupel{N, E, \mathcal{S}, \left(d_{i}\right)_{i\in N},
    \left(c_{e}\right)_{e\in E}}
  \end{equation*}
  \begin{equation*}
    \leadsto\mathcal{G} = \tupel{N, \mathcal{S}\times\mathcal{P}, \pi}
  \end{equation*}
  \vspace{-1cm}
  \begin{itemize}
    \item $\mathcal{P}=\times_{i\in N} P_{i}$ with
      $P_{i} = \mathbb{R}^{|E|}_{+}$: payment of player $i$
    \item $e\in E$ is bought under $S\in \mathcal{S}$ if
      $\sum_{i\in N}p_{i}^{e} \geq c_{e}(S)$
    \item $\pi_{i}(S) =
      \begin{cases}
        \sum_{e\in E}p_{i}^{e} &\textit{if }S_{i}\textit{ is bought}\\
        \infty &\textit{otherwise}
      \end{cases}$: {private cost of player $i$}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{A few simplifications}
  \begin{itemize}
    \item \emph{unweighted} games
    \item marginally non-increasing cost-function $c_{e}$:
      \begin{equation*}
        c_{e}(x + \delta) - c_{e}(x) \geq c_{e}(y + \delta) - c_{e}(y)
        \;\;x\leq y, \delta \in \mathbb{N}
      \end{equation*}
      \begin{center}
        \begin{tikzpicture}
          \begin{axis}[
              xlabel=$x$,
              ylabel=$c_{e}(x)$,
              smooth,
              width=5cm,
              height=3cm,
              ticks=none,
              no markers
            ]
              \addplot{ln(x)};
          \end{axis}
        \end{tikzpicture}
      \end{center}
    \item restrict players demands to a specific mathematical structure:
      Matroids
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Definition: Matroid}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{equation*}
      M = (E, B)
    \end{equation*}
    \vspace{-0.8cm}
    \begin{itemize}
      \item $E$: ground-set
      \item $B$: set of bases
      \item \emph{basis exchange property}:
        \vspace{-1cm}
        \begin{center}
          \begin{align*}
            \forall &X, Y\in B:\\
            &\forall x\in (X\setminus Y)\\
            &\exists y\in (Y\setminus X):\\
            &\;(X\setminus \set{x})\cup\{y\}\in B
          \end{align*}
        \end{center}
    \end{itemize}
    \column{0.5\textwidth}
    \uncover<2->{Spanning trees:}
    \begin{center}
      \uncover<2->{\resizebox{0.8\textwidth}{!}{\input{tikz/graphp1}}\\[0.5cm]}
      \uncover<3->{\resizebox{0.4\textwidth}{!}{\input{tikz/graphp1S1}}}
      \uncover<4->{\resizebox{0.4\textwidth}{!}{\input{tikz/graphp1S2}}}
      \uncover<5->{\resizebox{0.4\textwidth}{!}{\input{tikz/graphp1S3}}}
    \end{center}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Operations on Matroids $\mathcal{M} = \tupel{E, \mathcal{B}}$}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{center}
      Contraction:
    \end{center}
    \vspace{-0.6cm}
    \begin{align*}
      &\mathcal{M}/e = \tupel{E\setminus\set{e}, \mathcal{B}/e}\text{ with}\\
      &\mathcal{B}/e = \set{B\subseteq (E\setminus\set{e})\middle| B\cup\set{e}\in\mathcal{B}}
    \end{align*}
    \column<4->{0.5\textwidth}
    \begin{center}
      Deletion:
    \end{center}
    \vspace{-0.6cm}
    \begin{align*}
      &\mathcal{M}\setminus e = \tupel{E\setminus\set{e},\mathcal{B}\setminus e}\text{ with}\\
      &\mathcal{B}\setminus e = \set{B\subseteq E\setminus\set{e}\middle|B\in\mathcal{B}}
    \end{align*}
  \end{columns}
  \begin{center}
    \begin{tikzpicture}
      \uncover<2->{\node (G) {\input{tikz/graphp1}};
      \node[right=1 of G] (leadsto) {\Huge$\leadsto$};}
      \uncover<3>{\node[right=1 of leadsto] (G') {\input{tikz/graphp1Cont}};}
      \uncover<5>{\node[right=1 of leadsto] (G'') {\input{tikz/graphp1Del}};}
    \end{tikzpicture}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Cuts in Matroids}
  \begin{columns}
    \column{0.7\textwidth}
    $\mathcal{M} = \tupel{E, \mathcal{B}}$
    \begin{itemize}
      \item \emph{Cut}: inclusion-wise minimal $C\subseteq E$ such that
        $E\setminus C$ does not contain a basis of $\mathcal{M}$
      \item[$\leadsto$] $\mathcal{C}\tupel{\mathcal{M}} =
        \set{C\subseteq E\middle| C\text{ is a cut in } \mathcal{M}}$
    \end{itemize}
    \column{0.3\textwidth}
    \resizebox{0.8\textwidth}{!}{\input{tikz/graphp1}}
  \end{columns}
  \vspace{0.7cm}
  \begin{columns}
    \column{0.33\textwidth}
    \uncover<2->{\resizebox{0.7\textwidth}{!}{\input{tikz/graphp1Cut1}}\\[0.3cm]}
    \uncover<3->{\resizebox{0.7\textwidth}{!}{\input{tikz/graphp1Cut2}}}
    \column{0.33\textwidth}
    \uncover<4->{\resizebox{0.7\textwidth}{!}{\input{tikz/graphp1Cut3}}}\\[0.3cm]
    \uncover<5->{\resizebox{0.7\textwidth}{!}{\input{tikz/graphp1Cut4}}}
    \column{0.33\textwidth}
    \uncover<6->{\resizebox{0.7\textwidth}{!}{\input{tikz/graphp1Cut5}}}
  \end{columns}
\end{frame}

\section{Algorithm: PNE in Unweighted Matroid Games}
\subsection{Algorithm}
\begin{frame}
  \frametitle{Algorithm}
  \begin{tabular}{|lll|}
    \hline
    (i) & Init: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item \parbox{0.7\textwidth}{every player: empty working basis, complete ground-set, complete set of valid result bases}
          \item every resource: cost for next load, current load
        \end{itemize}
      }\\\hline
    (ii) & Find: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item inclusion-min (necessary cost)-max cut $C$
          \item minimal cost resource $e$ in $C$
          \item owner $i$ of $C$
        \end{itemize}
      }\\\hline
    (iii) & Pick: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item $i$: add $e$ to working basis, contract matroid to $e$
          \item $e$: update marginal cost and increase load
          \item all player: delete elements $C\setminus\set{e}$ from matroid
        \end{itemize}
      }\\\hline
    (iv) & Break: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item exists player with incomplete basis: goto (ii)
          \item otherwise: return chosen bases and payments
        \end{itemize}
      }\\\hline
  \end{tabular}
\end{frame}

\begin{frame}
  \frametitle{Example: Graph}
  \begin{columns}
    \column{0.5\textwidth}
    Player 1:\\
    \vspace{-0.2cm}
    \begin{center}
      \resizebox{0.6\textwidth}{!}{\input{tikz/graphp1labeled}}
    \end{center}
    \column{0.5\textwidth}
    Player 2:\\
    \vspace{-0.5cm}
    \begin{center}
      \resizebox{0.6\textwidth}{!}{\input{tikz/graphp2labeled}}
    \end{center}
  \end{columns}
  \begin{center}
    \resizebox{0.35\textwidth}{!}{\input{tikz/graphboth}}
  \end{center}
\end{frame}

\newlength{\cutscaling}
\setlength{\cutscaling}{0.17\textwidth}
\begin{frame}
  \frametitle{Example: Algorithm}
  \begin{columns}
    % left column:
    \column{0.3\textwidth}
    \resizebox{\textwidth}{!}{\alt<1-6>{\input{tikz/graphp1}}{\alt<7>{\input{tikz/graphp1Cont}}{\input{tikz/graphp1ContDel}}}}
    \begin{columns}
      \column{0.5\textwidth}
      \resizebox{\cutscaling}{!}{\alt<1-2>{\input{tikz/graphp1Cut1}}{\input{tikz/graphp1Cut1value}}}
      \resizebox{\cutscaling}{!}{\alt<1-3>{\input{tikz/graphp1Cut2}}{\input{tikz/graphp1Cut2value}}}
      \resizebox{\cutscaling}{!}{\alt<1-3>{\input{tikz/graphp1Cut3}}{\input{tikz/graphp1Cut3value}}}
      \column{0.5\textwidth}
      \resizebox{\cutscaling}{!}{\alt<1-3>{\input{tikz/graphp1Cut4}}{\input{tikz/graphp1Cut4value}}}
      \resizebox{\cutscaling}{!}{\alt<1-3>{\input{tikz/graphp1Cut5}}{\input{tikz/graphp1Cut5value}}}
    \end{columns}

    % middle column:
    \column{0.4\textwidth}
    \resizebox{\textwidth}{!}{\alt<1-8>{\input{tikz/graphbothCosts}}{\input{tikz/graphbothCostsUpd}}}
    \begin{center}
      \uncover<5->{\alt<5>{\resizebox{0.7\textwidth}{!}{\input{tikz/graphp1Cut3}}}{\resizebox{0.7\textwidth}{!}{\input{tikz/graphp1Cut3marked}}}}
    \end{center}

    % right column:
    \column{0.3\textwidth}
    \resizebox{\textwidth}{!}{\alt<1-7>{\input{tikz/graphp2}}{\input{tikz/graphp2Del}}}
    \begin{columns}
      \column{0.5\textwidth}
      \resizebox{\cutscaling}{!}{\alt<1-3>{\input{tikz/graphp2Cut1}}{\input{tikz/graphp2Cut1value}}}
      \resizebox{\cutscaling}{!}{\alt<1-3>{\input{tikz/graphp2Cut2}}{\input{tikz/graphp2Cut2value}}}
      \resizebox{\cutscaling}{!}{\alt<1>{\input{tikz/graphp2Cut3}}{\input{tikz/graphp2Cut3deleted}}}
      \column{0.5\textwidth}
      \resizebox{\cutscaling}{!}{\alt<1>{\input{tikz/graphp2Cut4}}{\input{tikz/graphp2Cut4deleted}}}
      \resizebox{\cutscaling}{!}{\alt<1>{\input{tikz/graphp2Cut5}}{\input{tikz/graphp2Cut5deleted}}}
      \resizebox{\cutscaling}{!}{\alt<1-3>{\input{tikz/graphp2Cut6}}{\input{tikz/graphp2Cut6value}}}
    \end{columns}
  \end{columns}
  \uncover<9>{ }
\end{frame}

\begin{frame}
  \frametitle{Soundness}
  \begin{itemize}
    \item Termination:
      \begin{itemize}
        \item inductively applying contraction operations lead to basis
        \item deletion cannot \enquote{loose} bases for other players since
          chosen cut is inclusion-wise minimal
      \end{itemize}
    \item Correctness
      \begin{itemize}
        \item Lemma: cost of the chosen element decreases monotonically in each
          iteration
        \item assume there is a beneficial alternative basis
          $\hat{B} = (B_{i}\setminus\set{g})\cup{f}$
          \begin{itemize}
            \item consider iteration $k$ in which $g$ is added to basis of $i$:
            \item if $f$ is in the corresonponding cut its cost never changes
              since it is deleted from every matroid $\rightarrow$ $f$ is not
              in the cut
            \item but $f$ and $g$ \enquote{open} same dimensionality, thus $f$
              is not left in contracted matroid of $i$
            \item $f$ was deleted before (in this iteration the cost of the
              chosen element was lower than for $f$ in contradiction to lemma)
          \end{itemize}
      \end{itemize}
  \end{itemize}
\end{frame}

\section{Weighted Matroid Games}
\begin{frame}
  \frametitle{Recall: Congestion Model}
  \begin{equation*}
    \mathcal{M} = \tupel{N, E, \mathcal{B}, \left(d_{i}\right)_{i\in N},
    \left(c_{e}\right)_{e\in E}}
  \end{equation*}
  \vspace{-1cm}
  \begin{itemize}
    \item before: $d_{i} = 1$ for all $i$
    \item now: $d_{i}\in\mathbb{N}_{>0}$ 
    \item auxiliary definitions:
      \begin{itemize}
        \item $N_{e}\tupel{B} = \set{i\in N\middle|e\in B_{i}}$
        \item for a fixed $B$: $\ex_{i}\tupel{e} =
          \set{f\in E\setminus\set{e}\middle|
          (B_{i}\setminus{e})\cup\set{f}\in\mathcal{B}_{i}}$: set of resources
          that can replace $e$ for player $i$
        \item $F = \bigcup_{i\in N}\bigcap_{B\in \mathcal{B}_{i}} B$:
          set of obligatory elements for at least one player
        \item $\Delta_{i}\tupel{B;e\rightarrow f} =
          c_{f}\tupel{\ell_{f}\tupel{B_{i}\cup\set{f}\setminus\set{e}}} -
          c_{f}\tupel{\ell_{f}\tupel{B}}$: marginal cost to buy resource $f$ if
          $i$ switches
        \item $\Delta_{i}^{e} = \min_{f\in\ex_{i}(e)}\set{\Delta_{i}(B;e\rightarrow f)}$:
          minimal marginal cost for exchanging $e$
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Characterisation of Pure Nash Equilibria}
  \begin{theorem}
    For a weighted matroid resource buying game
    $\tupel{N, E, \mathcal{B}, d, c}$ and $B\in\mathcal{B}$ the following
    holds: there is a payment vector $p$ such that $\tupel{B,p}$ is a
    PNE if and only if
    \begin{equation*}
      c_{e}(B) \leq \sum_{i\in N_{e}(B)}\Delta_{i}^{e}(B)
      \textit{ for all }e\in E\setminus F
    \end{equation*}
  \end{theorem}
  \begin{proof}
    if: $p_{i}^{e} =
      \begin{cases}
        \frac{\Delta_{i}^{e}(B)}{\sum_{j\in N_{e}(B)}\Delta_{j}^{e}(B)}\cdot
        c_{e}(B)&\text{ if } \sum_{j\in N_{e}(B)}\Delta_{j}^{e}(B) > 0\\
        0 &\text{ otherwise}
      \end{cases}$\\
      for all $e\in E\setminus F$ and $i\in N_{e}(B)$\\
    only if: for all $i$ and $e$: $p_{i}^{e} \leq \Delta_{i}^{e}(B)$
      and $c_{e}(B) = \sum_{i\in N_{e}(B)}p_{i}^{e}$
  \end{proof}
\end{frame}

\begin{frame}
  \frametitle{Existence of Pure Nash Equilibria}
  \begin{theorem}
    Every weighted matroid resource buying game with marginally non-increasing
    cost functions possesses a pure Nash equilibria.
  \end{theorem}
  \begin{itemize}
    \item socially optimal configuration profile: $B\in\mathcal{B}$ such that
      $\sum_{e\in E}c_{e}(\ell_{e}(B))$ is minimal
    \item for socially optimal configuration profile $B$ holds:
      \begin{equation*}
        c_{e}(B)\leq \sum_{i\in N_{e}(B)}\Delta_{i}^{e}(B)
        \text{ for all }e\in E\setminus F
      \end{equation*}
    \item $B$ is a pure Nash equilibrium
    \item (but finding socially optimal configuration profiles is NP-hard)
  \end{itemize}
\end{frame}

\section{Non-Matroid Strategy Spaces}
\begin{frame}
\frametitle{Consequences of using Non-Matroid Strategy Spaces}
  \begin{lemma}
    If $\mathcal{S}\subseteq 2^{E}$ is a non-matroid anti-chain, then there
    exist $X,Y\in\mathcal{S}$ and $\set{a,b,c}\subseteq X\cup Y$ such that each
    set $Z\in\mathcal{S}$ with $Z\subseteq(X\cup Y)\setminus\set{a}$ contains
    $b$ and $c$. (Especially, there is $S\in\mathcal{S}$ with
    $S\subseteq X\cup Y$ and $a\in S$ but $b,c\notin S$.)
  \end{lemma}
  \begin{theorem}
    For every non-matroid anti-chain $\mathcal{S}$ on a set of resource $E$,
    there exists a weighted two-player resource buying game
    $G = \tupel{\tilde{E}, \tupel{\mathcal{S}_{1} \times
    \mathcal{S}_{2}}\times \mathcal{P}, \pi}$ having marginally non-increasing
    cost functions, whose strategy spaces $\mathcal{S}_{1}$ and
    $\mathcal{S}_{2}$ are both isomophic to $\mathcal{S}$, so that $G$ does not
    possess a pure Nash equilibrium.
  \end{theorem}
\end{frame}

\begin{frame}
  \frametitle{Construction of games without PNE}

\end{frame}

\section{Conclusion}
\begin{frame}
\end{frame}

\end{document}