### 2017-04-07 paper.tex
polished

\documentclass{scrartcl}
\usepackage{pgfplots}
\usepackage{tikz}
\usepackage{csquotes}
\usepackage{framed}
\usepackage[hide]{todo}
\usepackage{amsmath}
\usepackage[framed]{ntheorem}
\usepackage[style=numeric]{biblatex}

\theoremstyle{nonumberplain}
\newframedtheorem{defi}{Definition}
\newframedtheorem{theo}{Theorem}
\newframedtheorem{lemma}{Lemma}

\addbibresource{bibliography.bib}

\newcommand{\tupel}[1]{\left(#1\right)}
\newcommand{\set}[1]{\left\{#1\right\}}

\DeclareMathOperator{\ex}{ex}

\title{Resource Buying Games}
\author{Christoph Welzel}

\begin{document}

\input{tikz/settings}
\maketitle
\section{Introduction}
In this paper we examine a certain class of games based on congestion models.
Intuitively speaking, we consider a set of resources and actors with individual
demands on this set of resources. We model this competing behaviour of actors
in terms of strategic games which are applicable for analysis regarding
Nash equilibria. Mainly, we are interested in the existence and computability
of pure Nash equilibria (PNE), i.e. a state where each actor has no interest
to unilaterally change her/his behaviour if all other players stick to their
strategy. We examine congestion models with different cost functions for
resources, different structures of resources each player wants to be available
and different demand of players on her/his resources. As main results we prove
the existence of PNE for
\begin{itemize}
  \item matroid games with marginally non-increasing cost functions
  \item congestion games with marginally non-decreasing cost functions
\end{itemize}
In terms of complexity we present an algorithm to compute a PNE for unweighted
(i.e. uniform demands for all players) matroid games with marginally
non-increasing cost functions. Additionally, we construct congestion games for
non-matroid structures with marginally non-increasing cost functions which
specifically do not allow for PNEs. This paper only presents results and
arguments from \cite{main}. Specifically, we focus on intuitions and thus
present proofs not in exhaustive detail but try to convey the central ideas.

We start by defining an appropiate congestion model and an associated
strategic game as well as by introducing matroids as combinatorial structures
to model desired configurations of resources for players. Following we present
an algorithm to compute PNEs in unweighted matroid games with marginally
non-increasing cost functions and its correctness proof.
Afterwards, we examine weighted matroid games with different cost functions
and concludingly, we study non-matroid configuration spaces.

\section{Congestion Model}
In this section, we introduce a structure to model resources and demands
of actors on resources. Therefore, we consider a set of players $N$ and a set
of resources $E$. Every player has a personal aim which can be achieved by
different resources. This is modelled for a player $i\in N$ as a family
$\mathcal{S}_{i}$ of sets $S_{i}\subseteq E$ of resources. Additionally, we
want to account for different demands of players on resources by fixing a
demand for every player $i$ as $d_{i}\in\mathbb{N}$. Finally, the cost of a
resource depends on the load the players put onto the resource modelled as
individual cost functions $c_{e}:\mathbb{N}\rightarrow \mathbb{R}_{+}$ for
every resource $e\in E$. As an example, consider different actors that build a
shared network infrastructure. The connections in the network are resources,
the actors are the players which have individual nodes they want to connect.
Different actors might need different bandwidth for connections, i.e. their
demands, and the cost for using connections depends on the used bandwidth.
Accordingly, we obtain a congestion model
\begin{equation*}
  \mathcal{M} = \tupel{N, E, \mathcal{S} = \times_{i\in N}\mathcal{S}_{i},
  (d_{i})_{i\in N}, (c_{e})_{e\in E}}
\end{equation*}
for which we introduce some additional definitions:
\begin{itemize}
  \item for a profile $S\in\mathcal{S}$ we define
    the demand on resource $e$ for $S$ as
    $\ell_{e}(S) = \sum_{i\in N:e\in S_{i}}d_{i}$
  \item we lift $c_{e}$ to the set $\mathcal{S}$ by setting
    $c_{e}(S) = c_{e}(\ell_{e}(S))$ for $S\in\mathcal{S}$
  \item we call $c_{e}$ \emph{marginally non-increasing} if
    $c_{e}(y + \delta) - c_{e}(y) \leq c_{e}(x + \delta) - c_{e}(x)$ holds
    for all $x,y,\delta\in N$ with $x\leq y$
  \item analogously $c_{e}$ is \emph{marginally non-decreasing} if
    $c_{e}(y + \delta) - c_{e}(y) \geq c_{e}(x + \delta) - c_{e}(x)$ holds
    for all $x,y,\delta\in N$ with $x\leq y$
\end{itemize}
Note that linear and constant cost functions are marginally non-increasing.

The cost that results from a profile $S\in\mathcal{S}$ is covered by all
players. This leads to a definition of resource buying game:
\begin{defi}
  For a congestion model $\mathcal{M} = \tupel{N, E, \mathcal{S},
  \tupel{d_{i}}_{i\in N}, \tupel{c_{e}}_{e\in E}}$ we define a
  \emph{resource buying game}
  \begin{equation*}
    G = \tupel{N, \mathcal{S}\times\mathcal{P}, \pi}
  \end{equation*}
  with $\mathcal{P} = \times_{i\in N}\mathcal{P}_{i}$ and
  $\mathcal{P}_{i} = \mathbb{R}_{+}^{|E|}$ (which is the contribution of each
  player to buy the demanded strategy profile).
  We denote the element from $\mathcal{P}_{i}$ for $e\in E$ with $p_{i}^{e}$
  and call $e\in E$ \emph{bought} under $\tupel{S, p}\in
  \mathcal{S}\times\mathcal{P}$ if $c_{e}(S)\leq \sum_{i\in N}p_{i}^{e}$. The
  function $\pi$ defines the payoff for every player for a strategy profile
  $\tupel{S, p}$ by
  \begin{equation*}
    \pi_{i}(S, p) = \begin{cases}
      \sum_{e\in E}p_{i}^{e} &\text{ if $S_{i}$ is bought under
      $\tupel{S, p}$ (i.e. all $e\in S_{i}$ are bought)}\\
      \infty &\text{ otherwise}
    \end{cases}
  \end{equation*}
\end{defi}
Thus, every player tries to minimise her/his personal contribution to the costs
but ensures that her/his demand is satisfied.
\section{Matroids}
In this section we present the combinatorial structure of matroids
\todo{why matroids?}
which we use as a generalisation of the concept of linear independence.
Matroids can be defined as a ground-set and the family of linearly
independent sets of this ground-set. This family has to satisfy three
properties, i.e.
\begin{enumerate}
  \item the empty set is linearly independent,
  \item all subsets of a linearly independent set are linearly independent and
  \item for two linearly independent sets we can find an element of the larger
    set that forms a linearly independent set when added to the smaller set.
\end{enumerate}
We use an equivalent definition of matroids which captures the idea of
bases of a vector space by giving the inclusion-wise maximal linear independent
sets (in analogy also called bases).
\begin{defi}
  The structure $\tupel{E, \mathcal{B}}$ with $\mathcal{B}\subseteq 2^{E}$ is
  a matroid if
  \begin{enumerate}
    \item $\mathcal{B}\neq\emptyset$
    \item let $B,B'\in\mathcal{B}$ and for every $x\in B$ there is $y\in B'$
      s. t. $(B\setminus\set{x})\cup\set{y}\in\mathcal{B}$
  \end{enumerate}
\end{defi}
In analogy to vector spaces we consider every element of a basis to open access
to a dimension. We illustrate this concept with the following example: given an
undirected connected graph $G = \tupel{V, E}$ we will argue that
$\tupel{E, \mathcal{S}}$ with $\mathcal{S}$ as edge-sets of spanning trees
in $G$ forms a matroid. In order to meet the requirement that $\mathcal{S}$
must not be empty we restrict $G$ to those graphs with at least two distinct
vertices. Furthermore, given two spanning trees $\tupel{V, S_{1}}$ and
$\tupel{V, S_{2}}$ of $G$ then removing an edge from $S_{1}$ yields a subgraph
of $G$ that is separated into two connected components. Since $S_{2}$ forms a
spanning tree as well there must be an edge connecting these components which
can be used to again obtain a spanning tree\footnote{see Figure
\ref{fig:spanning} for an illustration}.
\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \node (t) {\input{tikz/graphp1}};
      \node[below left=of t] {\input{tikz/graphp1S1}};
      \node[below =of t] {\input{tikz/graphp1S2}};
      \node[below right=of t] {\input{tikz/graphp1S1}};
    \end{tikzpicture}
  \end{center}
  \caption{Graph with its spanning trees.}
  \label{fig:spanning}
\end{figure}
\todo[done]{working iteratively towards a basis by kruskal's algorithm as
example to introduce cuts and contractions and deletions}
Consider the algorithm of Kruskal \cite{kruskal} which computes spanning trees
by iteratively adding edges to a result which do not close a cycle in the
subgraph induced by the edges in the result: every added
edge works towards a basis in $\mathcal{S}$ and opens a dimension by connecting
already computed components. When an edge is added the result of the algorithm
can only compute certain spanning trees, namely those which contain all edges
added up to this point. Fixing certain elements in matroids and therefore
restricting the set of \enquote{useable} bases is captured by the
\emph{contraction} operation on matroids.
\begin{defi}
  For a matroid $\mathcal{M} = \tupel{E, \mathcal{B}}$ and one $e\in E$ we call
  the matroid $\tupel{E\setminus\set{e}, \set{B\subseteq
  (E\setminus\set{e})\middle| B\cup\set{e}\in\mathcal{B}}}$ the
  \emph{contraction} of $e$ in $\mathcal{M}$, denoted by $\mathcal{M}/e$.
\end{defi}
In a similar manner we introduce the \emph{deletion} operation on matroids to
disregard all bases which use a certain element (in Kruskal's algorithm upon
adding an edge we can remove all other edges connecting the same components).
\begin{defi}
  For a matroid $\mathcal{M} = \tupel{E, \mathcal{B}}$ and one $e\in E$ we call
  the matroid $\tupel{E\setminus\set{e}, \set{B\subseteq
  (E\setminus\set{e})\middle| B\in\mathcal{B}}}$ the \emph{deletion} of $e$ in
  $\mathcal{M}$, denoted by $\mathcal{M}\setminus e$.
\end{defi}
In the context of graphs and the spanning tree matroids the contraction and
deletion operation corresponds to the contraction of two nodes connected by
an edge and the deletion of an edge respectively. Note that a deletion
operation is not necessarily well defined for all elements of the ground-set,
e.g. deletion of an edge in the graph which is a bridge yields
an empty set of bases in the spanning tree matroid and thus no valid matroid.
Analogously, any set of edges that, when removed, disconnects the graph
contains at least one element necessarily present in any basis. We call such
sets cuts\footnote{see Figure \ref{fig:cuts} for an illustration} and define
them as follows.  \begin{defi}
  Let $\mathcal{M} = \tupel{E, \mathcal{B}}$ be a matroid. A \emph{cut}
  $C\subseteq E$ is an inclusion-wise minimal set s.t. $E\setminus C$ does not
  contain any basis in $\mathcal{B}$.
\end{defi}
\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \node (1) {\input{tikz/graphp1Cut1}};
      \node[right=of 1] (2) {\input{tikz/graphp1Cut2}};
      \node[right=of 2] (3) {\input{tikz/graphp1Cut3}};
      \node[below=of 1] (4) {\input{tikz/graphp1Cut4}};
      \node[right=of 4] (5) {\input{tikz/graphp1Cut5}};
    \end{tikzpicture}
  \end{center}
  \caption{Cuts of the spanning tree matroid from the graph presented in Figure
    \ref{fig:spanning}.}
  \label{fig:cuts}
\end{figure}

\section{Unweighted Matroid Games with marginally non-increasing cost
functions}
We present an algorithm which computes pure Nash equilibria in unweighted
matroid games with marginally non-increasing cost functions. Firstly, an
unweighted matroid game is based on a congestion model of the form
$\tupel{N, E, \mathcal{B} = \times_{i\in N}\mathcal{B}_{i},
(d_{i})_{i\in N}, (c_{e})_{e\in E}}$ where for all $i\in N$ holds $d_{i} = 1$
(unweighted) and every $\mathcal{B}_{i}$ is the set of bases for some matroid
$\mathcal{M}_{i} = \tupel{E_{i}, \mathcal{B}_{i}}$. The set of resources is
given by the matroids of the players $E = \bigcup_{i\in N}E_{i}$ and all cost
functions $c_{e}$ are marginally non-increasing.

The algorithm we present in the following chooses a basis for every player by
iteratively adding one element from a cut and contracting the matroid to the
added element while deleting all other elements from the cut. Since cuts are
inclusion-wise minimal deleting all but one element from one cut leaves at
least one possible basis for the player available. For every player we maintain
the elements which are chosen for her/him up to the current state
(\emph{working basis}) and which elements are still available to build her
basis (\emph{ground-set}) and which bases she can still use. Initially the
working basis is empty and the ground-set and available bases are taken from
the matroid of the player. For every resource we maintain the cost the next
added load causes and how much load already is placed on the resource
(\emph{current load}). Initially, for every resource $e$ the current
load is zero and the next load causes $c_{e}(1)$ cost. We can divide the
algorithm into four steps as presented in Figure \ref{alg:main}.
\begin{figure}
  \begin{center}
    \begin{tabular}{|lll|}
      \hline
      (i) & Init: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
          \begin{itemize}
            \item \parbox{0.7\textwidth}{every player: empty working basis,
              complete ground-set, complete set of valid result bases}
            \item every resource: cost for next load, current load
          \end{itemize}
        }\\\hline
      (ii) & Find: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
          \begin{itemize}
            \item inclusion-min (necessary cost)-max cut $C$
            \item minimal cost resource $e$ in $C$
            \item owner $i$ of $C$
          \end{itemize}
        }\\\hline
      (iii) & Pick: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
          \begin{itemize}
            \item $i$: add $e$ to working basis, contract matroid to $e$
            \item $e$: update marginal cost and increase load
            \item all players: delete elements $C\setminus\set{e}$ from matroid
          \end{itemize}
        }\\\hline
      (iv) & Break: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
          \begin{itemize}
            \item if exists player with incomplete basis: goto (ii)
            \item otherwise: return chosen bases and payments
          \end{itemize}
        }\\\hline
    \end{tabular}
  \end{center}
  \caption{Algorithm to compute a PNE for unweighted matroid games with
    marginally non-increasing cost functions. Note that in step (ii) only
    inclusion-wise minimal cuts are considered (specifically the one for which
    minimal cost of all elements in the cut is maximal over all considered cuts
    is chosen). Restricting the selection to inclusion-wise minimal cuts
    ensures that removing all other elements preserves possible bases for
    \emph{all} players. Also the algorithm terminates in at most $|N|\cdot|E|$
    iterations (each iteration disgards at least one element for one player).
    For a more detailed pseudo-code representation and a description of how to
    compute the minimal cut and the minimal cost element in this cut we refer
    to \cite[Algorithm 1 and 3.3 A Subroutine to Detect $C^{\ast}, e^{\ast}$
    and $i^{\ast}$]{main}.}
  \label{alg:main}
\end{figure}
By induction on the number of iterations we can prove that the cost of the
chosen element in step (ii) always decreases. This is a key insight to prove
the correctness of the algorithm. Because, if we assume to have a beneficial
alternative to the computed basis for one player there must exist an element
with less cost. This alternative element cannot be part of the selected cut
since it would be removed for all players and thus keep its marginal cost
although only elements with higher costs are removed (by choice of the element
in the cut). Furthermore, the alternative element cannot be available to the
player for which the cut is selected since both elements open the
\enquote{same dimension} in contradiction to the choice of a cut. Thus, the
alternative element must already have been removed which implies less
cost in a previous step to the iteration where the original element was chosen.
This contradicts the monotonicity of costs of considered elements through the
iterations. {Thus, we get the following theorem:
\begin{theo}
  The presented algorithm computes a PNE.
\end{theo}
}

\section{Weighted Matroid Games}
Starting from a congestion model where every player aims to buy a basis of
a certain matroid and every player puts an individual demand upon the resources
she/he uses we derive a complete characterisation of PNEs as well as show that
every socially optimal solution induces a PNE. First of all, we give some
auxiliary definitions (with a profile $B\in\mathcal{B}$, a resource $e\in E$):
\begin{itemize}
  \item $N_{e}(B) = \set{i\in N\middle| e\in B_{i}}$ is the set of players
    which use resource $e$ in the profile $B$
  \item $\ex_{i}(e) = \set{f\in E\setminus\set{e}\middle|
    (B_{i}\setminus\set{e})\cup\set{f}\in\mathcal{B}_{i}}$ is the set of
    resources player $i\in N_{e}(B)$ could exchange $e$ with
  \item $F$ is the set of resources that is obligatory for any player, i.e.
    every element of $F$ is present in \emph{all} bases of one (or more) player
  \item $\Delta_{i}\tupel{B;e\rightarrow f}$ is the marginal cost player $i$
    has to pay for resource $f$ if he intends to use $f$ instead of $e$.
    We define the minimum over all resources $e$ can be exchanged with as
    $\Delta_{i}^{e}(B) = \min_{f\in\ex_{i}(e)}\Delta_{i}(B;e\rightarrow f)$
\end{itemize}
With these definitions we can formulate the following characterisation of
PNEs which holds for all non-negative and non-decreasing cost functions.
\begin{theo}
  For a weighted matroid resource buying game $\tupel{N, E, \mathcal{B}, d, c}$
  there is a payment vector $p$ s.t. $\tupel{B\in\mathcal{B}, p}$ is a PNE
  if and only if
  \begin{equation*}
    c_{e}(B)\leq\sum_{i\in N_{e}(B)}\Delta_{i}^{e}(B) \text{ for all } e\in
    E\setminus F
  \end{equation*}
\end{theo}
The \enquote{only if} direction is easily shown by the fact that the cost
of a resource equals the sum of the contributions of every player for this
resource and no player has an incentive to unilaterally change to another basis
of her/his matroid. While, for the \enquote{if} direction, we let for elements
in $F$ one player for whom this resource is obligatory pay for this resource
(since she/he needs this resource to be available she/he can not choose to
unilaterally change her/his payment). For all other elements we can define the
payments as follows:
\begin{equation*}
  p_{i}^{e} = \begin{cases}
    0 &\text{ if } \sum_{j\in N_{e}(B)}\Delta_{j}^{e}(B) = 0\\
    \underbrace{
      \frac{c_{e}(B)}{\sum_{j\in N_{e}(B)}\Delta_{j}^{e}(B)}
    }_{\leq 1}\cdot\Delta_{i}^{e}(B) &\text{ otherwise}
  \end{cases}
\end{equation*}
which yields $p_{i}^{e}\leq \Delta_{i}^{e}$ for all $e\in E\setminus F$
implying a PNE.

Using this we can derive a series of inequalities to show that every socially
optimal solution satisfies the given characterisation and hence:
\begin{theo}
  Every weighted matroid resource buying game with marginally non-increasing
  cost functions possesses a PNE.
\end{theo}
For completeness we state - without going into details - an analogous result
for non-decreasing cost functions. Again starting from a socially optimal
configuration we can derive a pricing vector (called marginal cost pricing)
which forms a PNE with the given configuration. Noteably, this result is
also true for non-matroid strategy spaces, yielding:
\begin{theo}
  Let $G$ be a weighted resource buying game with non-decreasing marginal costs
  and let $S^{\ast}$ be a socially optimal solution. Then, marginal cost
  pricing induces a PNE.
\end{theo}

\section{Non-Matroid Strategy Spaces}
Finally, we consider non-matroid strategy spaces. Specifically,
we present a construction to obtain a weighted two player
resource buying game with marginally non-increasing cost functions which does
not posses a PNE starting from a non-matroid anti-chain $S$. This implies that
matroids are exactly the structures to consider using marginally non-increasing
cost functions since other structures cannot guarantee the existence of PNEs.
Thus, we prove the following theorem:
\begin{theo}
  For every non-matroid anti-chain $\mathcal{S}$ (i.e. a family of sets which
  are not ordered by $\subseteq$) on a set of resources $E$, there exists a
  weighted two-player resource buying game $G = \tupel{\tilde{E},
    (\mathcal{S}_{1}\times \mathcal{S}_{2})\times\mathcal{P},\pi}$
  having marginally non-increasing cost functions, whose strategy spaces
  $\mathcal{S}_{1}$ and $\mathcal{S}_{2}$ are both isomorphic to $\mathcal{S}$,
  so that $G$ does not possess a PNE.
\end{theo}
We start with an anti-chain $\mathcal{S}\subseteq 2^{E}$ which
specifically does \emph{not} satisfy the matroid properties. Hence
$\mathcal{S}$ contains two sets $X, Y$ s.t. there is one element in $X$ for
which we cannot find a single element in $Y$ to obtain an element from
$\mathcal{S}$ again. Intuitively, the dimension of one element in $X$ needs
two elements from $Y$ to become available. This can be formally described
by the following lemma:
\begin{lemma}
  If $\mathcal{S}\subseteq 2^{E}$ is a non-matroid anti-chain, then there exist
  $X,Y\in\mathcal{S}$ and $\set{a,b,c}\subseteq X\cup Y$ such that each
  $Z\in\mathcal{S}$ with $Z\subseteq(X\cup Y)\setminus\set{a}$ contains both,
  $b$ and $c$.
\end{lemma}
Especially, we can guarantee that there are two elements in $\mathcal{S}$ s.t.
one contains $a$ (but not $b$ and $c$) and the other $b$ and $c$ (but not $a$).
In the following we construct a game which does not possess a PNE. The
constructed game is illustrated in Figure \ref{fig:game} for better
understanding.
\begin{figure}
  \input{tikz/gamediagram}
  \caption{Illustration of constructed game which does not possess a PNE.}
  \label{fig:game}
\end{figure}
We consider two isomorphic copies of $\mathcal{S}$:
$\mathcal{S}_{1},\mathcal{S}_{2}$ and by previous lemma there are
$X_{i}, Y_{i}\in\mathcal{S}_{i}$ for $i = 1,2$ as described above. We
interleave the ground-sets of $\mathcal{S}_{1}$ and $\mathcal{S}_{2}$ by
precisely three elements $x,y,z$ which we identify differently in both copies.
Namely, we set $x = a_{1} = b_{2}$, $y = b_{1} = a_{2}$, and
$z = c_{1} = c_{2}$. The set of resources for $G$ is now the union of both
ground-sets w.r.t. their intersection which is $\set{x,y,z}$. Furthermore, we
fix the cost functions of all elements in
$\tilde{E}\setminus(X_{1}\cup X_{2}\cup Y_{1}\cup Y_{2})$ to a high constant
$M$ which ensures that both players only play within $X_{i},
Y_{i}$ (for $i = 1,2$ respectively). In this \enquote{sub-arena} we set the
cost functions of all elements other than $\set{x,y,z}$ to $0$ regardless of
the demand. And concludingly, we set $c_{x}(t) = t$, $c_{y}(t) = 5\frac{1}{2}$,
and $c_{z}(t) = 4$ and an individual demand of the players of $d_{1} = 5$ and
$d_{2} = 4$. Thus, each player either contributes to pay for her/his $a$ or
to $b$ and $c$. Assuming there is a PNE
$\tupel{Z^{\ast} = \tupel{Z_{1}^{\ast}, Z_{2}^{\ast}},
p^{\ast} = \tupel{p_{1}^{\ast}, p_{2}^{\ast}}}$ we can conclude by the choice
of $\set{x,y,z}$ that
\begin{equation*}
  x\notin Z_{1}^{\ast}\Rightarrow \set{y,z}\subseteq Z_{1}^{\ast} \text{ and }
  y\notin Z_{2}^{\ast}\Rightarrow \set{x,z}\subseteq Z_{2}^{\ast}
\end{equation*}
By examining the restricted environment of the cost inducing elements
$\set{x,y,z}$ and the conclusion that one player either only contributes to $x$
or $y$ and $z$ and the other only to $y$ or $x$ and $z$ one can conclude that
neither $x\in Z_{1}^{\ast}$ nor $x\notin Z_{1}^{\ast}$ which yields a
contradiction to the existence of a PNE and concludes the argument on the
theorem proposed at the beginning of the section.

\section{Conclusion}
Among other restrictions we studied unweighted congestion games with marginally
non-increasing cost functions. For these games matroids are precisely the
configuration spaces to guarantee PNEs as well as allow for efficient
computation of one PNE. For weighted matroid games with marginally
non-increasing cost functions the existence of PNEs can be assured as well as
for weighted congestion games with marginally non-decreasing cost functions.
Hereby the structure of strategy spaces is unrestricted.
Concludingly, marginally non-increasing cost functions demand matroid strategy
spaces as combinatorial structures which capture e.g. spanning trees, linear
independence, etc. On the other hand, marginally non-decreasing cost functions
might induce a certain stability in a market by ensuring the existence of PNEs.

\printbibliography

\todos

\end{document}